---
title: "DAFI Paper"
author: "James Millington"
date: "2021-06-14"
output: 
  github_document:
    toc: true
---

# Overview
Analysis of DAFI Fire Use variables by Anthropogenic Fire Regime. Also distinguishing Broadcast vs Cropland fires. 

Load packages, data and supporting functions. Code not shown in rendered document (i.e. html, pdf, etc) but can be viewed in original .rmd file.
```{r load-packages, message=F, echo=F, warning=F}

library(tidyverse)
library(ggplot2)
library(viridisLite)
library(maps)
library(xlsx)
library(openxlsx)
library(devtools)
library(readxl)
library(scales)

vir_10 <- viridis_pal(option='inferno')(10)

```

```{r load-scripts}

source("manip-script.r", local = knitr::knit_global())
source("plots-script.r", local = knitr::knit_global())
```

```{r summarise-function, include=F}

load.db()

```


# 3. Results

## 3.1 Database Summary

```{r}
ns <- n_distinct(record_info$`Study.ID`)
ncs <- n_distinct(record_info$`Case.Study.ID`)

pubs <- record_info %>%
  group_by(Study.Type) %>%
  summarise(n = n()) %>%
  mutate(perc = 100 * n / sum(n))
pubs
aca <- round(filter(pubs, Study.Type=='Academic')$perc,0)
govngo <- round(sum(filter(pubs, Study.Type=='NGO' | Study.Type=='Government')$perc),0)

datsrc <- record_info %>%
  mutate(Data.Source.s.=ifelse(grepl(',',Data.Source.s.), 'Mixed', as.character(Data.Source.s.))) %>%
  mutate(Data.Source.s.=ifelse(Data.Source.s.=='Review', 'Literature review', as.character(Data.Source.s.))) %>%
  group_by(Data.Source.s.) %>%
  summarise(n = n()) %>%
  mutate(perc = 100 * n / sum(n))


datsrc
src.fs <- round(filter(datsrc, Data.Source.s.=='Primary')$perc,0)
src.idr <- round(filter(datsrc, Data.Source.s.=='Secondary')$perc,0)
src.rs <- round(filter(datsrc, Data.Source.s.=='Remote sensing')$perc,0)
src.lr <- round(filter(datsrc, Data.Source.s.=='Literature review')$perc,0)
src.oth <- round(filter(datsrc, Data.Source.s.=='Other (see notes)')$perc,0)
src.mix <- round(filter(datsrc, Data.Source.s.=='Mixed')$perc,0)

#sum(datsrc$perc)
```
In this initial data collection, data from `r ns` studies were collated from across `r ncs` human fire-related case studies. 

Data were overwhelmingly from academic publications (`r aca`% of case studies) but were also from reports produced by governments and NGOs (`r govngo`%). 

Case studies used data exclusively from Field Studies (`r src.fs`% of case studies), Institutional Data Repositories (`r src.idr`%), Literature Review (`r src.lr`%), Remote Sensing (`r src.rs`%) or a combination (`src.mix`%) of these data source types. The remaining case studies (~`r src.oth`%) used expert elicitation, media reports, archival research and other reports as sources.

When these data sources are mapped spatially (Figure 1) we find a prevalence of case studies using institutional data in Europe and North America versus a dominance of field studies in Asia and Africa.  


### Figure 1

```{r message=F, warning=F}
srcmap <- map.behaviour.ras('Data.Source.s.', ras.function='mode')
srcmap

afrmap <- map.behaviour.ras('Anthropogenic fire regime', ras.function='mode')
afrmap

fumap <- map.behaviour.ras('Fire purpose', ras.function='mode')
fumap

```


```{r}
#number of case studies
ncs.rfire <- n_distinct(rep_fire_use$`Case Study ID`)

#count of incomplete case studies by variable
incsvar.rfire <- rep_fire_use %>%
  dplyr::select(-Notes, -Calculations) %>%
  group_by(`Case Study ID`) %>%
  summarise_each(funs(sum(is.na(.)))) %>%
  mutate_all(funs(ifelse(.>1,1,.))) %>%
  summarise_all(sum)

#count of missing values across all variables
mval.rfire <- incsvar.rfire %>%
  dplyr::select(-`Case Study ID`) %>%
  sum()

#total number of possible values
tval.rfire <- ncs.rfire * dim(incsvar.rfire)[2] - 1   #-1 to ignore the Case Study ID

#% missing values
tval.rfire.perc <- round(100* mval.rfire/ tval.rfire,0)


#count of incomplete case studies by variable for QUANTITATIVE vars only
incsvar.rfire.q <- rep_fire_use %>%
  dplyr::select_if(names(.)=="Case Study ID" | sapply(., is.numeric)) %>%
  dplyr::select(-`Study Year`) %>%
  group_by(`Case Study ID`) %>%
  summarise_each(funs(sum(is.na(.)))) %>%
  mutate_all(funs(ifelse(.>1,1,.))) %>%
  summarise_all(sum)

#count of missing values across all variables
mval.rfire.q <- incsvar.rfire.q %>%
  dplyr::select(-`Case Study ID`) %>%
  sum()

#total number of possible values
tval.rfire.q <- ncs.rfire * dim(incsvar.rfire.q)[2] - 1   #-1 to ignore the Case Study ID

#% missing values
tval.rfire.q.perc <- round(100* mval.rfire.q / tval.rfire.q,0)

```


Reflecting the fragmented nature of the anthropogenic fire literature, no case study contained data on all variables and data across all DAFI fields are sparse. For example, for reported fire use data, `r tval.rfire.perc`% of values were missing, rising to `r tval.rfire.q.perc`% when only quantitative variables are considered. 

```{r}

lu.AFR <- land_use %>% 
  rename(AFR = `Anthropogenic fire regime`) %>%
  count(AFR, .drop=FALSE) %>%
  drop_na()

cs.AFR <- sum(lu.AFR$n)

filter(lu.AFR, AFR=="Pre-industrial")$n

```

Furthermore, we were able to define an AFR for only `r cs.AFR` of the `ncs` case studies (Pre-industrial `r filter(lu.AFR, AFR=="Pre-industrial")$n`, Transition `r filter(lu.AFR, AFR=="Transition")$n`, Industrial `r filter(lu.AFR, AFR=="Industrial")$n`, Post-industrial `r filter(lu.AFR, AFR=="Post-industrial")$n`). The incompleteness of the database means that the number of values used in analyses below varies depending on which aspects of anthropogenic fire are being examined. 


```{r}

afrsrc <- merge(record_info, land_use, by.x= 'Case.Study.ID', by.y = 'Case Study ID')

afrsrc <- afrsrc %>% 
  mutate(Data.Source.s.=ifelse(grepl(',',Data.Source.s.), 'Mixed', as.character(Data.Source.s.))) %>%
  mutate(Data.Source.s.=ifelse(Data.Source.s.=='Review', 'Literature review', as.character(Data.Source.s.))) %>%
  dplyr::select(Data.Source.s., `Anthropogenic fire regime`)

table(afrsrc)

afrsrc.s <- summary(table(afrsrc))
afrsrc.p <- round(prop.table(table(afrsrc),2),2)

```

The spatial distribution of AFRs (Figure 1b) indicates a similar distribution to data sources, with a prevalence of Industrial and Post-Industrial regimes in Europe and North America versus a dominance of Transition and Pre-Industrial regimes in Asia and Africa. A Chi-square test for the association between AFR and data source indicates non-randomness (Chi-sq = `round(afrsrc.s$statistic,0)`, df = `afrsrc.s$parameter`, p < 0.001) and Primary sources dominate case studies of Pre-industrial (`100*max(afrsrc.p[,'Pre-industrial'])`% of case studies) and Transition (`100*max(afrsrc.p[,'Transition'])`%) regimes, whereas Secondary sources are the most frequent data sources for case studies of Industrial (`100*max(afrsrc.p[,'Industrial'])`%) and Post-industrial (`100*max(afrsrc.p[,'Post-industrial'])`%) regimes.   

## Fire Purpose

```{r}
rep_FUS <- rep_fire_use
est_FUS <- est_fire_use

rep_FUS$`Fire purpose` <- Simplify.purpose(rep_fire_use$`Fire purpose`, AFT = rep_fire_use$AFT)
est_FUS$`Fire purpose` <- Simplify.purpose(est_fire_use$`Fire purpose`, AFT = est_fire_use$AFT)

#summarise grouped (sum) by case study, AFR and fire use
fs.cs.afr <- plyr::rbind.fill(lapply(colnames(rep_fire_use)[c(13:20)],
                                     function (x){summarise.behaviour(temp.rep_fire_use=rep_FUS,temp.est_fire_use=est_FUS,
                                                                      type = "Fire", behaviour = x,
                                                                      #sum_multi = c('Case Study ID', 'Anthropogenic fire regime','Fire purpose','Study Year'),
                                                                      grouping = c('Case Study ID', 'Anthropogenic fire regime','Fire purpose'),
                                                                      inc.Absence = F, escape.rm = T)}))

#summarise grouped (sum) by case study, AFR and fire use
ba.cs.afr <- plyr::rbind.fill(lapply(colnames(rep_fire_use)[c(21:26)],
                                     function (x){summarise.behaviour(temp.rep_fire_use=rep_FUS,temp.est_fire_use=rep_FUS,
                                                                      type = "Fire", behaviour = x,
                                                                      #sum_multi = c('Case Study ID', 'Anthropogenic fire regime','Fire purpose'),
                                                                      grouping = c('Case Study ID', 'Anthropogenic fire regime','Fire purpose'),
                                                                      inc.Absence = F, escape.rm = T)}))

#summarise grouped (sum) by case study, AFR and fire use
rp.cs.afr <- plyr::rbind.fill(lapply(colnames(rep_fire_use)[27],
                                     function (x){summarise.behaviour(temp.rep_fire_use=rep_FUS,temp.est_fire_use=rep_FUS,
                                                                      type = "Fire", behaviour = x,
                                                                      #sum_multi = c('Case Study ID', 'Anthropogenic fire regime','Fire purpose'),
                                                                      grouping = c('Case Study ID', 'Anthropogenic fire regime','Fire purpose'),
                                                                      inc.Absence = F, escape.rm = T)}))
```


### Table 2

```{r}

#DAFI Records %
fu.count <- rep_FUS %>%
  drop_na(`Fire purpose`) %>%
  tally()

fu.perc <- rep_FUS %>%
  drop_na(`Fire purpose`) %>%
  group_by(`Fire purpose`) %>%
  summarise(n=n()) %>%
  mutate(fuperc = 100* n / as.numeric(fu.count))

fu.perc  

#% that the seven main fire use tyoe compose
fu.contrib <- sum(fu.perc$fuperc) - fu.perc$fuperc[fu.perc$`Fire purpose` == 'Other']

#Mean Size (ha)  #intended fire size mean
fs.fp <- plyr::rbind.fill(lapply(colnames(rep_fire_use)[17],
                                     function (x){summarise.behaviour(temp.rep_fire_use=rep_FUS,temp.est_fire_use=est_FUS,
                                                                      type = "Fire", behaviour = x,
                                                                      grouping = c('Fire purpose'),
                                                                      inc.Absence = F, escape.rm = T)}))
fu.sizemn <- fs.fp %>%
  dplyr::select(`Fire purpose`, `Combined.stat`)

fu.sizemn 


###
#Mean burned area (% LS)  - unclear how are these being calclated?

#summarise grouped (sum) by case study, fire use
ba.cs <- plyr::rbind.fill(lapply(colnames(rep_fire_use)[c(21:26)],
                                     function (x){summarise.behaviour(temp.rep_fire_use=rep_FUS,temp.est_fire_use=rep_FUS,
                                                                      type = "Fire", behaviour = x,
                                                                      #sum_multi = c('Case Study ID', 'Anthropogenic fire regime','Fire purpose'),
                                                                      grouping = c('Fire purpose'),
                                                                      inc.Absence = F, escape.rm = T)}))


#return period (years)
rp.cs <- plyr::rbind.fill(lapply(colnames(rep_fire_use)[27],
                                     function (x){summarise.behaviour(temp.rep_fire_use=rep_FUS,temp.est_fire_use=rep_FUS,
                                                                      type = "Fire", behaviour = x,
                                                                      #sum_multi = c('Case Study ID', 'Anthropogenic fire regime','Fire purpose'),
                                                                      grouping = c('Fire purpose'),
                                                                      inc.Absence = F, escape.rm = T)}))
rp.cs 


#escaped - unclear how these are being calculated 
typ.count <- rep_FUS %>%
  drop_na(`Fire type`) %>%
  drop_na(`Fire purpose`) %>%
  group_by(`Fire purpose`) %>%
  filter(grepl('Human',`Fire type`)) %>%
  summarise(n=n())

esc.count <- rep_FUS %>%
  drop_na(`Fire type`) %>%
  drop_na(`Fire purpose`) %>%
  group_by(`Fire purpose`) %>%
  filter(`Fire type` == 'Human, escaped') %>%
  summarise(n=n())

esc.perc <- esc.count %>%
  mutate(escperc = 100 * n / typ.count$n)

esc.perc

```

After such similar types were combined (see online appendix B!), seven dominant fire uses emerged (Table 2), each with more than 100 instances in the database and accounting for `r fu.contrib`% of fire instance records.

### Figure 2

```{r}
a <- filter(rep_fire_use, `Presence / Absence` == 'Presence')
b <- filter(est_fire_use, `Presence / Absence` == 'Presence')

fig2 <- bar.purpose.regime(a, b, bartype='fill')
ggsave('/home/james/OneDrive/Research/Papers/InProgress/Perkins_DAFI/figures/Figure2.png', plot=fig2, height=12, width=14, units='cm', dpi=300)
bar.purpose.regime(rep_fire_use, est_fire_use, bartype='dodge')

fig2

fig2all <- bar.purpose.regime(rep_fire_use, est_fire_use, bartype='fill')
fig2all
```



```{r}
# final paper
# 
rep_FUS <- rep_fire_use
est_FUS <- est_fire_use

rep_FUS$`Fire purpose` <- Simplify.purpose(rep_fire_use$`Fire purpose`, AFT = rep_fire_use$AFT)
est_FUS$`Fire purpose` <- Simplify.purpose(est_fire_use$`Fire purpose`, AFT = est_fire_use$AFT)



#summarise grouped (sum) by case study, AFR and fire use
fs.cs.afr <- plyr::rbind.fill(lapply(colnames(rep_fire_use)[c(13:20)],
                                     function (x){summarise.behaviour(temp.rep_fire_use=rep_FUS,temp.est_fire_use=est_FUS,
                                                                      type = "Fire", behaviour = x,
                                                                      #sum_multi = c('Case Study ID', 'Anthropogenic fire regime','Fire purpose','Study Year'),
                                                                      grouping = c('Case Study ID', 'Anthropogenic fire regime','Fire purpose'),
                                                                      inc.Absence = F, escape.rm = T)}))
```

**Data in DAFI show that deliberate anthropogenic fires occur (where present) at a median and mean rate of 0.075 and 1.83 km-2 year-1 respectively. Highly dense anthropogenic fires are found in very fertile agricultural areas such as the Mekong River Delta, where data in DAFI suggest 96 fires km-2 Year-1. At the lower end of the scale, pasture fires in DAFI are reported at 0.09 fires km-2 Year-1. Similarly, fire return intervals are typically short, with median of 3 and mean of 7.14 years. **


### Figure 3


```{r}
fs_mean_kde <- hist.fire.regime(dat.temp=fs.cs.afr,metric = 'fire size mean', bin_width = -1, actual = F, log_scale=T, regime=F, vertical=T, scale_limits = c(0.1, 10000)) 
#ggsave('/home/james/OneDrive/Research/Papers/InProgress/Perkins_DAFI/figures/fs_mean_kde.pdf', plot=fs_mean_kde, height=14, width=9, units='cm')
fs_mean_kde

fs_mean_box <- box.FS.regime(fs.cs.afr, metric = 'mean', boxtype='box', 
                   estimated = T, reported = T, log_scale = T, cropland = T, scale_limits = c(0.1, 10000))
#ggsave('/home/james/OneDrive/Research/Papers/InProgress/Perkins_DAFI/figures/fs_mean_box.pdf', plot=fs_mean_box, height=14, width=16, units='cm')
fs_mean_box 
```


```{r}
ba_ha_box <- box.BA.regime(dat.temp=ba.cs.afr,metric = 'burned area (ha)', boxtype='box', 
              log_scale = T, cropland = T, scale_limits = c(0.01, 1000000000)) 
#ggsave('/home/james/OneDrive/Research/Papers/InProgress/Perkins_DAFI/figures/ba_ha_box.pdf', plot=ba_ha_box, height=14, width=16, units='cm')
ba_ha_box

ba_ha_kde <- hist.fire.regime(dat.temp=ba.cs.afr,metric = 'burned area (ha)', bin_width = -1, log_scale=T, regime=F, vertical=T,  scale_limits = c(0.01, 1000000000)) 
#ggsave('/home/james/OneDrive/Research/Papers/InProgress/Perkins_DAFI/figures/ba_ha_kde.pdf', plot=ba_ha_kde, height=14, width=9, units='cm')
ba_ha_kde

```



## 3.3 Fire Suppression

### Figure 4

```{r}

#suppression at case study level
#
#only count CaseStudy once for each AFR-Suppression-Effort combination
#don't use pivot_longer as that results in double-counting

cleanEffort <- function(supTbl, supLabel){
  
  rsupTbl <- supTbl %>%
    setNames(c("Case","AFR","Effort")) %>%
    drop_na() %>%
    filter(Effort != 'ND') %>%
    mutate(Effort = recode(Effort, `0`='None',`1`='Limited',`2`='Moderate',`3`='Intensive')) %>%
    group_by_all %>%
    summarise(n= n()) %>%
    dplyr::select(-n) %>%
    mutate(Suppression = supLabel)
  
  return(rsupTbl)
  
}

sup.c <- cleanEffort(dplyr::select(suppression, 1,2,9),'Control')
sup.p <- cleanEffort(dplyr::select(suppression, 1,2,11),'Prevention')
sup.e <- cleanEffort(dplyr::select(suppression, 1,2,13),'Extinction')

c.sup <- rbind(sup.c, sup.p, sup.e)  

c.sup.c <- c.sup %>% 
  ungroup() %>%
  mutate(Effort = factor(Effort, levels = c("None","Limited","Moderate","Intensive"))) %>%
  mutate(AFR = factor(AFR, levels = c("Pre-industrial","Transition","Industrial","Post-industrial"))) %>% 
  mutate(Suppression = factor(Suppression,levels = c("Control","Prevention","Extinction"))) %>%
  count(AFR, Suppression, Effort,.drop=FALSE)


#paper figure!
p3 <- c.sup.c %>%
  ggplot(aes(x = AFR, y=n)) +
  geom_bar(aes(fill=Effort),position='dodge',stat="identity") +
  facet_grid(Suppression~.) +
  scale_fill_viridis_d(option = 'inferno') +
  xlab("Anthropogenic Fire Regime") +
  ylab("Count")




p4 <- c.sup.c %>%
  ggplot(aes(x = AFR, y=n)) +
  geom_bar(aes(fill=Effort),position='fill',stat="identity") +
  facet_grid(Suppression~.) +
  scale_fill_viridis_d(option = 'inferno') +
  xlab("Anthropogenic Fire Regime") +
  ylab("Count")

p3
p4

```


```{r}
c.sup.c %>%
 filter(Effort != "None") %>%
 group_by(AFR, Suppression) %>%
 summarise(sum = sum(n))

c.sup.c %>%
 group_by(AFR, Effort) %>%
 summarise(sum = sum(n))

filter(c.sup, AFR=='Post-industrial', Effort=="Intensive", Suppression=="Extinction")

```

## 3.4 Fire Policy

```{r}

pol.sum <- plyr::rbind.fill(lapply(colnames(policy)[c(10,14,19)],
                                     function (x){summarise.behaviour(type = "Policy", behaviour = x,
                                                                      grouping = c('Anthropogenic fire regime','Direct Fire policy ID'),
                                                                      inc.Absence = F, escape.rm = T)}))
pol.sum <-pol.sum[!is.na(pol.sum$`Anthropogenic fire regime`),]
pol.sum <- pol.sum %>% 
    count(`Anthropogenic fire regime`, Behaviour,.drop=FALSE)

pol.sum$`Anthropogenic fire regime` <- factor(pol.sum$`Anthropogenic fire regime`, levels = c("Pre-industrial","Transition","Industrial","Post-industrial"))

p7 <- pol.sum %>%
  ggplot(aes(x = `Anthropogenic fire regime`, y=n)) +
  geom_bar(aes(fill=Behaviour),position='dodge',stat="identity") +
  scale_fill_viridis_d(option = 'inferno') +
  xlab("Anthropogenic Fire Regime") +
  ylab("Count")

p7

```

```{r}

#policy at lowest level of data ('Direct Fire Policy ID')
pol <- policy %>%
  dplyr::select(2,4,10,14,19)  %>%
  setNames(c("AFR","Type","Bans","Restrictions","Incentives")) %>%
  filter(Type == 'Human, deliberate') %>% 
  mutate(Bans = case_when(grepl("Yes", Bans, ignore.case=TRUE) ~ "Yes")) %>%
  mutate(Restrictions = case_when(grepl("Yes", Restrictions, ignore.case=TRUE) ~ "Yes")) %>%
  mutate(Incentives = case_when(grepl("Yes", Incentives, ignore.case=TRUE) ~ "Yes")) %>%
  dplyr::select(-Type)

pol <-pivot_longer(pol, cols=c("Bans":"Incentives"), names_to="Policy")

#set factors, drop NAs and count
pol <- pol %>%
  mutate(Policy = factor(Policy, levels=c("Incentives","Restrictions","Bans"))) %>%
  mutate(AFR = factor(AFR, levels = c("Pre-industrial","Transition","Industrial","Post-industrial"))) %>% 
  drop_na(value) %>%
  drop_na(AFR) %>%
  count(AFR, Policy,.drop=FALSE)

pol.plot <- pol %>%
  ggplot(aes(x = AFR, y=n)) +
  geom_bar(aes(fill=Policy),position='dodge',stat="identity") +
  scale_fill_viridis_d(option = 'inferno') +
  xlab("Anthropogenic Fire Regime") +
  ylab("Count")


```

```{r}

#policy at lowest level of data ('Direct Fire Policy ID')
pol.pid <- policy %>%
  dplyr::select(2,4,10:17,19:22) %>%
  rename('AFR' = 1) %>%
  #rename('Health restriction' = 'Human health restriction') %>%
  #rename('Health ban' = 'Human health ban') %>%
  #filter(`Fire type` == 'Human, deliberate') %>% 
  mutate_at(vars(-1,-2), funs(case_when(grepl("Yes", ., ignore.case=TRUE) ~ "Yes"))) %>%
  dplyr::select(-`Fire type`)

#create 'Other' Rationale for when Type = Yes but individual rationale is not specified
pol.pid <- pol.pid %>% 
  rename('Unknown incentives' = 'Incentives') %>%
  rename('Unknown restriction' = 'Fire restricted') %>%
  rename('Unknown ban' = 'Fire banned') %>%
  mutate(`Unknown incentives` = if_else(`Unknown incentives` == "Yes" & is.na(`Economic incentives`) &
                                          is.na(`Environmental incentives`) & is.na(`Health incentives`), "Yes", NA_character_)) %>%
  mutate(`Unknown restriction` = if_else(`Unknown restriction` == "Yes" & is.na(`Economic restriction`) &
                                          is.na(`Environmental restriction`) & is.na(`Health restriction`), "Yes", NA_character_)) %>%
  mutate(`Unknown ban` = if_else(`Unknown ban` == "Yes" & is.na(`Economic ban`) &
                                          is.na(`Environmental ban`) & is.na(`Health ban`), "Yes", NA_character_))

#go long
pol.pid <-pivot_longer(pol.pid, cols=c(-1), names_to="Rationale-Type")

#create new columns for Rationale and Type then set factors and levels
pol.pid <- pol.pid %>%
  separate(`Rationale-Type`, c("Rationale", "Type"), sep=" ") %>%
  mutate(Type = recode_factor(Type, incentives='Incentive',restriction='Restriction',ban='Ban')) %>%
  mutate(Rationale = factor(Rationale, levels=c("Environmental","Economic","Health","Unknown"))) %>%
  mutate(AFR = factor(AFR, levels = c("Pre-industrial","Transition","Industrial","Post-industrial")))

#drop NAs and count         
pol.pid <- pol.pid %>% 
  drop_na(value) %>%
  drop_na(AFR) %>%
  count(AFR, Type, Rationale,.drop=FALSE)

#plot
p.pol.pid <- pol.pid %>%
  ggplot(aes(x = AFR, y=n)) +
  geom_bar(aes(fill=Type),position='dodge',stat="identity") +
  facet_grid(Rationale~.) +
  scale_fill_viridis_d(option = 'inferno') +
  xlab("Anthropogenic Fire Regime") +
  ylab("Count")

p.pol.pid
```

```{r}

#paper figure!

#policy at case study level 
pol.cs <- policy %>%
  dplyr::select(1,2,4,10:17,19:22) %>%
  rename('Case'=1,'AFR' = 2) %>%
  #rename('Health restriction' = 'Human health restriction') %>%
  #rename('Health ban' = 'Human health ban') %>%
  #filter(`Fire type` == 'Human, deliberate') %>% 
  mutate_at(vars(-1,-2,-3), funs(case_when(grepl("Yes", ., ignore.case=TRUE) ~ "Yes"))) %>%
  dplyr::select(-`Fire type`)

#create 'Other' Rationale for when Type = Yes but individual rationale is not specified
pol.cs <- pol.cs %>% 
  rename('Unknown incentives' = 'Incentives') %>%
  rename('Unknown restriction' = 'Fire restricted') %>%
  rename('Unknown ban' = 'Fire banned') %>%
  mutate(`Unknown incentives` = if_else(`Unknown incentives` == "Yes" & is.na(`Economic incentives`) &
                                          is.na(`Environmental incentives`) & is.na(`Health incentives`), "Yes", NA_character_)) %>%
  mutate(`Unknown restriction` = if_else(`Unknown restriction` == "Yes" & is.na(`Economic restriction`) &
                                          is.na(`Environmental restriction`) & is.na(`Health restriction`), "Yes", NA_character_)) %>%
  mutate(`Unknown ban` = if_else(`Unknown ban` == "Yes" & is.na(`Economic ban`) &
                                          is.na(`Environmental ban`) & is.na(`Health ban`), "Yes", NA_character_))

pol.cs <- pol.cs %>%
  group_by_all %>%
  summarise(n= n()) %>%
  ungroup()

#go long
pol.cs <-pivot_longer(pol.cs, cols=c(`Unknown incentives`:`Health ban`), names_to="Rationale-Type")

#create new columns for Rationale and Type then set factors and levels
pol.cs <- pol.cs %>%
  separate(`Rationale-Type`, c("Rationale", "Type"), sep=" ") %>%
  mutate(Type = recode_factor(Type, incentives='Incentive',restriction='Restriction',ban='Ban')) %>%
  mutate(Rationale = factor(Rationale, levels=c("Environmental","Economic","Health","Unknown"))) %>%
  mutate(AFR = factor(AFR, levels = c("Pre-industrial","Transition","Industrial","Post-industrial")))

#drop NAs and count         
pol.cs <- pol.cs %>% 
  drop_na(value) %>%
  drop_na(AFR) %>%
  count(AFR, Type, Rationale,.drop=FALSE)

#plot
p.pol.cs <- pol.cs %>%
  ggplot(aes(x = AFR, y=n)) +
  geom_bar(aes(fill=Type),position='dodge',stat="identity") +
  facet_grid(Rationale~.) +
  scale_fill_viridis_d(option = 'inferno') +
  xlab("Anthropogenic Fire Regime") +
  ylab("Count")

#plot stacked
s.pol.cs <- pol.cs %>%
  ggplot(aes(x = AFR, y=n)) +
  geom_bar(aes(fill=Type),position='fill',stat="identity") +
  facet_grid(Rationale~.) +
  scale_fill_viridis_d(option = 'inferno') +
  xlab("Anthropogenic Fire Regime") +
  ylab("Count")

p.pol.cs
s.pol.cs
```

```{r}


pol.cs %>%
  group_by(Type) %>%
  summarise(sum = sum(n))

```
# 4. Discussion

## 4.1 Improving the quality of anthropogenic fire data

```{r}
ba.rep <- ba.cs.afr %>%
      drop_na(`Anthropogenic fire regime`) %>%
      filter(Combined.stat > 0.01) %>%
      filter(grepl('burned area (ha)',Behaviour,fixed=TRUE)) 

fs.rep  <- fs.cs.afr %>%
      drop_na(`Anthropogenic fire regime`) %>%
      filter(Combined.stat > 0.01) %>%
      filter(grepl('mean',Behaviour,fixed=TRUE))

ba.rep <- round(100 * length(ba.rep[,1]) / n_distinct(record_info$`Case.Study.ID`),1)
fs.rep <- round(100 * length(fs.rep[,1]) / n_distinct(record_info$`Case.Study.ID`),1)
```

we find that in case studies, burned area metrics are reported less frequently than fire size metrics (`r ba.rep`% of case studies vs `r fs.rep`%). 

##  4.2 Modelling and observing anthropogenic fire regimes 
### Figure 6

```{r}
#https://datascience.blog.wzb.eu/2019/04/30/zooming-in-on-maps-with-sf-and-ggplot2/
##https://ggplot2.tidyverse.org/reference/ggsf.html
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
worldmap <- ne_countries(scale = 'medium', type = 'map_units',returnclass = 'sf')


dat <- right_join(record_info, est_fire_use, by=c("Case.Study.ID"='Case Study ID'))

dat$Latitude <- as.numeric(dat$Latitude)
dat$Longitude<- as.numeric(dat$Longitude) 

fsdat <- dat %>% 
  filter(grepl('Crop',`Fire purpose`)) %>%
  drop_na(`Fire season start`) %>%
  mutate(`Fire season start` = factor(`Fire season start`, levels = month.name)) %>%
  mutate(`Fire quarter` = factor(case_when(
    `Fire season start` == 'December' | `Fire season start` == 'January' | `Fire season start` == 'February' ~ 'DJF',
    `Fire season start` == 'March' | `Fire season start` == 'April' | `Fire season start` == 'May' ~ 'MAM',
    `Fire season start` == 'June' | `Fire season start` == 'July' | `Fire season start` == 'August' ~ 'JJA',
    `Fire season start` == 'September' | `Fire season start` == 'October' | `Fire season start` == 'November' ~ 'SON'), 
    levels= c('DJF','MAM','JJA','SON')))

shapes = c(21,23) 
inset_x <- c(70, 95)
inset_y <- c(20, 33)

#hack described here: https://github.com/tidyverse/ggplot2/issues/3909#issuecomment-602253881
myrect <- 
  list(
    cbind(
      c(inset_x[1], inset_x[2], inset_x[2], inset_x[1], inset_x[1]), 
      c(inset_y[1], inset_y[1], inset_y[2], inset_y[2], inset_y[1])
    )
  ) %>%
  st_polygon() %>%
  st_sfc(crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

w <- ggplot() +
  geom_sf(data = worldmap, fill="beige", colour='grey',size=0.1) +
  geom_sf(data = myrect,
  #geom_rect(xmin = 50, xmax = 70, ymin = 50, ymax = 70, 
     fill = NA, colour = "red", size = 0.5) +
  coord_sf(xlim = c(-170, 181), ylim = c(-60, 85), expand = FALSE) +
  geom_point(data=fsdat,aes(x = Longitude, y = Latitude, shape = `Fire purpose`, fill= `Fire quarter`),
             colour='black', size = 0.75, stroke=0.1, alpha=0.9) + 
  theme(
    panel.background = element_rect(fill = "azure",colour = "black",size = 0.5, linetype = "solid"),
    panel.grid.major = element_line(linetype='blank'),
    panel.grid.minor = element_line(linetype='blank'),
    legend.key = element_rect(fill = "white", colour = "white")
  ) +
  scale_fill_viridis_d(option='inferno') + 
  guides(fill=guide_legend(override.aes=list(shape=22))) +
  scale_shape_manual(values=shapes) 


i <- ggplot() + 
  geom_sf(data = worldmap, fill='beige', colour='grey') + 
  coord_sf(xlim = inset_x, ylim = inset_y, expand = FALSE) +
  geom_point(data=fsdat, aes(x = Longitude, y = Latitude, shape = `Fire purpose`, fill= `Fire quarter`),
             colour='black', size = 2, stroke=0.5, alpha=0.9) +
    theme(
    panel.background = element_rect(fill = "azure",colour = "red",size = 0.5, linetype = "solid"),
    panel.grid.major = element_line(linetype='blank'),
    panel.grid.minor = element_line(linetype='blank'),
    legend.key = element_rect(fill = "white", colour = "white")
  ) +
  scale_fill_viridis_d(option='inferno') + 
  guides(fill=guide_legend(override.aes=list(shape=22))) +
  scale_shape_manual(values=shapes) 
```

```{r}
w
i

```



```{r}

count21ha <- rep_FUS %>%
  filter(`Actual fire size mean (ha)` > 21) %>%
  filter(`Presence / Absence` == 'Presence') %>%
  filter(`Fire type` == 'Human, deliberate') %>%
  nrow(.)
  
count0ha <- rep_FUS %>%
  filter(`Actual fire size mean (ha)` > 0) %>%
  filter(`Presence / Absence` == 'Presence') %>%
  filter(`Fire type` == 'Human, deliberate') %>%
  nrow(.)  

perc21 <- 100 - (100 * as.numeric(count21ha) / as.numeric(count0ha))

count100ha <- rep_FUS %>%
  filter(`Actual fire size mean (ha)` > 100) %>%
  filter(`Presence / Absence` == 'Presence') %>%
  filter(`Fire type` == 'Human, deliberate') %>%
  nrow(.)
  
perc100 <- 100 - (100 * as.numeric(count100ha) / as.numeric(count0ha))

```

We find that `r perc21`% of records in DAFI for mean size of deliberately started fires are <21 ha, suggesting many anthropogenic fires will not be detected. 

Our results from DAFI support this; `r perc100`% of mean fire size records for deliberately started fires are <100ha, and cropland fires in pre-industrial and transition regimes are generally smaller compared to fires ignited to burn across landscapes more broadly (Figure 3x!).
