---
title: "DAFI Paper"
author: "James Millington"
date: "2022-06-29"
output: 
  github_document:
    toc: true
---

# Overview
Analysis of DAFI Fire Use variables by Anthropogenic Fire Regime. Also distinguishing Broadcast vs Cropland fires. 

All derived values are shown in context of the text in which they are used.

## Setup
Load packages, data and supporting functions. Code not shown in rendered document (i.e. html, pdf, etc) but can be viewed in original .rmd file. Session Info shown for package versions.  
```{r load-packages, message=F, echo=F, warning=F}

library(tidyverse)
library(ggplot2)
library(viridisLite)
library(maps)
library(xlsx)
library(openxlsx)
library(devtools)
library(readxl)
library(scales)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)

vir_10 <- viridis_pal(option='inferno')(10)
```

```{r load-scripts}
source("manip-script.r", local = knitr::knit_global())
source("plots-script.r", local = knitr::knit_global())
```

```{r sessionInfo}
sessionInfo()
```

```{r load.db, include=F}
load.db()
```

# 2. Materials and Methods

## 2.3 Database Summary and Analysis
```{r}
ns <- n_distinct(record_info$`Study.ID`)
ncs <- n_distinct(record_info$`Case.Study.ID`)

pubs <- record_info %>%
  group_by(Study.Type) %>%
  summarise(n = n()) %>%
  mutate(perc = 100 * n / sum(n))
pubs
aca <- round(filter(pubs, Study.Type=='Academic')$perc,0)
govngo <- round(sum(filter(pubs, Study.Type=='NGO' | Study.Type=='Government')$perc),0)

datsrc <- record_info %>%
  mutate(Data.Source.s.=ifelse(grepl(',',Data.Source.s.), 'Mixed', as.character(Data.Source.s.))) %>%
  mutate(Data.Source.s.=ifelse(Data.Source.s.=='Review', 'Literature review', as.character(Data.Source.s.))) %>%
  group_by(Data.Source.s.) %>%
  summarise(n = n()) %>%
  mutate(perc = 100 * n / sum(n))

datsrc
src.fs <- round(filter(datsrc, Data.Source.s.=='Primary')$perc,0)
src.idr <- round(filter(datsrc, Data.Source.s.=='Secondary')$perc,0)
src.rs <- round(filter(datsrc, Data.Source.s.=='Remote sensing')$perc,0)
src.lr <- round(filter(datsrc, Data.Source.s.=='Literature review')$perc,0)
src.oth <- round(filter(datsrc, Data.Source.s.=='Other (see notes)')$perc,0)
src.mix <- round(filter(datsrc, Data.Source.s.=='Mixed')$perc,0)

#sum(datsrc$perc)
```
At the time of publication DAFI contains data from `r ncs` human fire‐related case studies collated from `r ns` sources.

Data were overwhelmingly from academic publications (`r aca`% of case studies) but were also from reports produced by governments and NGOs (`r govngo`%). 

Case studies used data exclusively from Field Studies (`r src.fs`% of case studies), Institutional Data Repositories (`r src.idr`%), Literature Review (`r src.lr`%), Remote Sensing (`r src.rs`%) or a combination (`r src.mix`%) of these data source types. The remaining case studies (~`r src.oth`%) used expert elicitation, media reports, archival research and other reports as sources.



```{r warning=FALSE}
#number of case studies
ncs.rfire <- n_distinct(rep_fire_use$`Case Study ID`)

#count of incomplete case studies by variable
incsvar.rfire <- rep_fire_use %>%
  dplyr::select(-Notes, -Calculations) %>%
  group_by(`Case Study ID`) %>%
  summarise_each(funs(sum(is.na(.)))) %>%
  mutate_all(funs(ifelse(.>1,1,.))) %>%
  summarise_all(sum)

#count of missing values across all variables
mval.rfire <- incsvar.rfire %>%
  dplyr::select(-`Case Study ID`) %>%
  sum()

#total number of possible values
tval.rfire <- ncs.rfire * dim(incsvar.rfire)[2] - 1   #-1 to ignore the Case Study ID

#% missing values
tval.rfire.perc <- round(100* mval.rfire/ tval.rfire,0)

#count of incomplete case studies by variable for QUANTITATIVE vars only
incsvar.rfire.q <- rep_fire_use %>%
  dplyr::select_if(names(.)=="Case Study ID" | sapply(., is.numeric)) %>%
  dplyr::select(-`Study Year`) %>%
  group_by(`Case Study ID`) %>%
  summarise_each(funs(sum(is.na(.)))) %>%
  mutate_all(funs(ifelse(.>1,1,.))) %>%
  summarise_all(sum)

#count of missing values across all variables
mval.rfire.q <- incsvar.rfire.q %>%
  dplyr::select(-`Case Study ID`) %>%
  sum()

#total number of possible values
tval.rfire.q <- ncs.rfire * dim(incsvar.rfire.q)[2] - 1   #-1 to ignore the Case Study ID

#% missing values
tval.rfire.q.perc <- round(100* mval.rfire.q / tval.rfire.q,0)
```

Reflecting the fragmented nature of the anthropogenic fire literature, no case study contained data on all variables and data across all DAFI fields are sparse. For example, for reported fire use data, `r tval.rfire.perc`% of values were missing, rising to `r tval.rfire.q.perc`% when only quantitative variables are considered. 

```{r}
lu.AFR <- land_use %>% 
  rename(AFR = `Anthropogenic fire regime`) %>%
  count(AFR, .drop=FALSE) %>%
  drop_na()

cs.AFR <- sum(lu.AFR$n)
```

Furthermore, we were able to define an AFR for only `r cs.AFR` of the `r ncs` case studies (Pre-industrial `r filter(lu.AFR, AFR=="Pre-industrial")$n`, Transition `r filter(lu.AFR, AFR=="Transition")$n`, Industrial `r filter(lu.AFR, AFR=="Industrial")$n`, Post-industrial `r filter(lu.AFR, AFR=="Post-industrial")$n`) as land use information is not always provided or available. The incompleteness of the database means that the number of values used in analyses below varies depending on which aspects of anthropogenic fire are being examined. 

```{r}
afrsrc <- merge(record_info, land_use, by.x= 'Case.Study.ID', by.y = 'Case Study ID')

afrsrc <- afrsrc %>% 
  mutate(Data.Source.s.=ifelse(grepl(',',Data.Source.s.), 'Mixed', as.character(Data.Source.s.))) %>%
  mutate(Data.Source.s.=ifelse(Data.Source.s.=='Review', 'Literature review', as.character(Data.Source.s.))) %>%
  dplyr::select(Data.Source.s., `Anthropogenic fire regime`)

table(afrsrc)

afrsrc.s <- summary(table(afrsrc))
afrsrc.p <- round(prop.table(table(afrsrc),2),2)
```

# 3. Results

## 3.1 Geographic Distribuion

When DAFI data sources are mapped spatially (Figure 1) we find a prevalence of case studies using institutional data in Europe and North America versus a dominance of field studies in Asia and Africa. 

### Figure 1
```{r message=F, warning=F}
srcmap <- map.behaviour.ras('Data.Source.s.', ras.function='mode')
srcmap

afrmap <- map.behaviour.ras('Anthropogenic fire regime', ras.function='mode')
afrmap

fumap <- map.behaviour.ras('Fire purpose', ras.function='mode')
fumap
```
The spatial distribution of AFRs (Figure 1b) indicates a similar distribution to data sources, with a prevalence of Industrial and Post-Industrial regimes in Europe and North America versus a dominance of Transition and Pre-Industrial regimes in Asia and Africa. A Chi-square test for the association between AFR and data source indicates non-randomness (Chi-sq = `r round(afrsrc.s$statistic,0)`, df = `r afrsrc.s$parameter`, p < 0.001) and Primary sources dominate case studies of Pre-industrial (`r 100*max(afrsrc.p[,'Pre-industrial'])`% of case studies) and Transition (`r 100*max(afrsrc.p[,'Transition'])`%) regimes, whereas Secondary sources are the most frequent data sources for case studies of Industrial (`r 100*max(afrsrc.p[,'Industrial'])`%) and Post-industrial (`r 100*max(afrsrc.p[,'Post-industrial'])`%) regimes.   

## 3.2. Fire Use
### 3.2.1. Fire Purpose
```{r}
rep_FUS <- rep_fire_use
est_FUS <- est_fire_use

rep_FUS$`Fire purpose` <- Simplify.purpose(rep_fire_use$`Fire purpose`, AFT = rep_fire_use$AFT)
est_FUS$`Fire purpose` <- Simplify.purpose(est_fire_use$`Fire purpose`, AFT = est_fire_use$AFT)

#summarise grouped (sum) by case study, AFR and fire use
fs.cs.afr <- plyr::rbind.fill(lapply(colnames(rep_fire_use)[c(13:20)],
                                     function (x){summarise.behaviour(temp.rep_fire_use=rep_FUS,temp.est_fire_use=est_FUS,
                                                                      type = "Fire", behaviour = x,
                                                                      sum_multi = F,
                                                                      grouping = c('Case Study ID', 'Anthropogenic fire regime','Fire purpose'),
                                                                      inc.Absence = F, escape.rm = T)}))

#summarise grouped (sum) by case study, AFR and fire use
ba.cs.afr <- plyr::rbind.fill(lapply(colnames(rep_fire_use)[c(21:26)],
                                     function (x){summarise.behaviour(temp.rep_fire_use=rep_FUS,temp.est_fire_use=rep_FUS,
                                                                      type = "Fire", behaviour = x,
                                                                      sum_multi=F,
                                                                      grouping = c('Case Study ID', 'Anthropogenic fire regime','Fire purpose'),
                                                                      inc.Absence = F, escape.rm = T)}))
```

```{r}
#DAFI Records %
fu.count <- rep_FUS %>%
  drop_na(`Fire purpose`) %>%
  tally()

fu.perc <- rep_FUS %>%
  drop_na(`Fire purpose`) %>%
  group_by(`Fire purpose`) %>%
  summarise(n=n()) %>%
  mutate(fuperc = 100* n / as.numeric(fu.count))

#% that the seven main fire use tyoe compose
fu.contrib <- round(sum(fu.perc$fuperc) - fu.perc$fuperc[fu.perc$`Fire purpose` == 'Other'],0)
```

Overall, 20 anthropogenic fire uses were identified during literature review, but many were closely related (e.g., ‘pasture renewal’ and ‘rangeland management’). After such similar types were combined (see Supplementary Materials), seven dominant fire uses emerged (Table 2) each with more than 100 instances in the database and accounting for `r fu.contrib`% of fire instance records.

### Table 2

#### DAFI records
```{r}
fu.perc
```

#### Mean size (ha)
```{r}
#summarise
fs.fp <- plyr::rbind.fill(lapply(colnames(rep_fire_use)[17],
                                     function (x){summarise.behaviour(temp.rep_fire_use=rep_FUS,temp.est_fire_use=est_FUS,
                                                                      type = "Fire", behaviour = x,
                                                                      sum_multi = F,
                                                                      grouping = c('Fire purpose'),
                                                                      inc.Absence = F, escape.rm = T)}))
fu.sizemn <- fs.fp %>%
  dplyr::select(`Fire purpose`, `Combined.stat`)

fu.sizemn 
```

#### Mean burned area (% LS)
```{r}
#summarise
ba.fp        <- plyr::rbind.fill(lapply(colnames(rep_FUS)[c(23:27)],
                                        function (x){summarise.behaviour(temp.rep_fire_use=rep_FUS,temp.est_fire_use=est_FUS,
                                                                         type = "Fire", sum_multi = FALSE,
                                                                         behaviour = x, grouping = c('Fire purpose'), 
                                                                         escape.rm = T)}))
ba.fp$Intended <- ifelse(grepl('intended', tolower(ba.fp$Behaviour)), 'Intended', 'Actual')

ba.fp.i <- ba.fp %>%
  filter(grepl('land cover', Behaviour)) %>%
  filter(Intended == 'Intended') %>%
  rbind(c('Hunter gatherer',rep(0,6),'Intended burned area % (land cover)','Intended')) %>%
  mutate(Combined.stat = as.numeric(Combined.stat), 
         Combined.N = as.numeric(Combined.N)) %>%
  arrange(`Fire purpose`)
  
ba.fp.a <- ba.fp %>%
  filter(grepl('land cover', Behaviour)) %>%
  filter(Intended == 'Actual') %>%
  mutate(Combined.stat = as.numeric(Combined.stat), 
         Combined.N = as.numeric(Combined.N)) %>%
  drop_na(`Fire purpose`) %>%
  arrange(`Fire purpose`)

ba.fp.lc <- round(((ba.fp.i$Combined.stat * ba.fp.i$Combined.N) + (ba.fp.a$Combined.stat * ba.fp.a$Combined.N)) / (ba.fp.i$Combined.N + ba.fp.a$Combined.N),1)

ba.fp.lc <- cbind.data.frame(ba.fp.i$`Fire purpose`, ba.fp.lc)

ba.fp.lc
```

#### Return period (years)
```{r}
#summarise
rp.fp <- plyr::rbind.fill(lapply(colnames(rep_FUS)[27],
                                     function (x){summarise.behaviour(temp.rep_fire_use=rep_FUS,temp.est_fire_use=est_FUS,
                                                                      type = "Fire", behaviour = x,
                                                                      sum_multi = F,
                                                                      grouping = c('Fire purpose'),
                                                                      inc.Absence = F, escape.rm = T)}))
dplyr::select(rp.fp, `Fire purpose`, Combined.stat) 
```

#### Escaped %

For fires uses with complete data 
```{r}
#summarise
igs <-  plyr::rbind.fill(lapply(colnames(rep_FUS)[c(11:12)],
                                  function (x){summarise.behaviour(temp.rep_fire_use=rep_FUS,temp.est_fire_use=est_FUS,
                                                                   type = "Fire", sum_multi =  FALSE,
                                                                   behaviour = x, grouping = c('Fire type', 'Fire purpose'), 
                                                                   inc.Absence = F, escape.rm = F)}))

igs.e <- igs %>% 
  filter(`Fire type` == 'Human, escaped' & !is.na(`Fire purpose`) & grepl('land cover',Behaviour)) %>%
  filter(`Fire purpose` != 'Hunter gatherer') %>%
  mutate(Combined.stat = as.numeric(Combined.stat), 
         Combined.N = as.numeric(Combined.N)) %>%
  arrange(`Fire purpose`)

igs.d <- igs %>% 
  filter(`Fire type` == 'Human, deliberate' & !is.na(`Fire purpose`) & grepl('land cover',Behaviour)) %>%
  filter(!(`Fire purpose` %in% c('Arson','Pyrome management','Vegetation clearance'))) %>%
  mutate(Combined.stat = as.numeric(Combined.stat), 
         Combined.N = as.numeric(Combined.N)) %>%
  arrange(`Fire purpose`) 

esc.n <- rep_FUS %>% 
  filter(`Fire type` == 'Human, escaped'  & !is.na(`Fire purpose`)) %>%
  group_by_at(.vars = c('Fire purpose', 'Fire type')) %>%
  summarise(count = n()) %>%
  filter(`Fire purpose` %in% c('Crop field preparation', 'Crop residue burning', 'Pasture management')) %>%
  arrange(`Fire purpose`)

del.n <-  rep_FUS %>% 
  filter(`Fire type` == 'Human, deliberate'  & !is.na(`Fire purpose`)) %>%
  group_by_at(.vars = c('Fire purpose', 'Fire type')) %>%
  summarise(count = n()) %>%
  filter(`Fire purpose` %in% c('Crop field preparation', 'Crop residue burning', 'Pasture management')) %>%
  arrange(`Fire purpose`)

escaped <- 
round(100*(esc.n$count / (esc.n$count + del.n$count)) * (igs.e$Combined.stat / (igs.e$Combined.stat + igs.d$Combined.stat)),2)

escaped <- cbind.data.frame(del.n$`Fire purpose`, escaped)
escaped
```

For less complete fire uses, use global (non-cropland) means with overall counts
```{r}
#summarise
igs.o <-  plyr::rbind.fill(lapply(colnames(rep_FUS)[c(11:12)],
                                  function (x){summarise.behaviour(temp.rep_fire_use=rep_FUS,temp.est_fire_use=est_FUS,
                                                                   type = "Fire", sum_multi =  FALSE,
                                                                   behaviour = x, grouping = c('Fire type', 'Fire purpose'), 
                                                                   escape.rm = F)}))

igs.o.d <- igs.o %>% filter(`Fire type` == 'Human, deliberate' & !is.na(`Fire purpose`)) %>%
  filter(grepl('total', Behaviour)) %>%
  filter(!(`Fire purpose` %in% c('Crop field preparation', 'Crop residue burning'))) %>%
  filter(`Fire purpose` != 'Other')

denom <- sum(igs.o.d$Combined.stat*igs.o.d$Combined.N) / sum(igs.o.d$Combined.N)


igs.o.e <- igs.o %>% filter(`Fire type` == 'Human, escaped' & !is.na(`Fire purpose`)) %>%
  filter(!(`Fire purpose` %in% c('Crop field preparation', 'Crop residue burning'))) %>%
  filter(grepl('total', Behaviour)) %>%
  filter(`Fire purpose` != 'Other')

numer <-sum(igs.o.e$Combined.stat*sum(igs.o.e$Combined.N)) / sum(igs.o.e$Combined.N)

esc.n.g <- rep_FUS%>% 
  filter(`Fire type` == 'Human, escaped'  & !is.na(`Fire purpose`)) %>%
  group_by_at(.vars = c('Fire purpose', 'Fire type')) %>%
  filter(!(`Fire purpose` %in% c('Crop field preparation', 'Crop residue burning', 'Pasture management', 'Other', 'Arson'))) %>%
  summarise(count = n())

del.n.g <- rep_FUS%>% 
  filter(`Fire type` == 'Human, deliberate'  & !is.na(`Fire purpose`)) %>%
  group_by_at(.vars = c('Fire purpose', 'Fire type')) %>%
  filter(!(`Fire purpose` %in% c('Crop field preparation', 'Crop residue burning', 'Pasture management', 'Other', 'Arson'))) %>%
  summarise(count = n())

escaped.g <- 
round(100*(esc.n.g$count / (esc.n.g$count + del.n.g$count)) * (numer / denom),2)

escaped.g <- cbind.data.frame(del.n.g$`Fire purpose`, escaped.g)
escaped.g
```

### Figure 2
```{r}
a <- filter(rep_fire_use, `Presence / Absence` == 'Presence')
b <- filter(est_fire_use, `Presence / Absence` == 'Presence')

fig2 <- bar.purpose.regime(a, b, bartype='fill')
fig2 <- fig2 + theme(axis.text.x=element_text(color = "black", size=9),
                     axis.text.y=element_text(color = "black", size=9)) +
  labs(fill = "Fire Use", x = 'Anthropogenic Fire Regime')
#ggsave(fig2, filename='Figure2_600.png', width=14, height=10, units='cm', dpi=600)

fig2_alt <- bar.purpose.regime(rep_fire_use, est_fire_use, bartype='dodge')

fig2
fig2_alt
```
### 3.2.2. Physical Characteristics

```{r}
#summarise
igs <- plyr::rbind.fill(lapply(colnames(rep_FUS)[c(11:12)],
                      function (x){summarise.behaviour(temp.rep_fire_use=rep_FUS,temp.est_fire_use=est_FUS,
                                                       type = "Fire", behaviour = x, 
                                                       grouping = c('Case Study ID'), escape.rm = T)}))

igs.med <- round(median(igs$Combined.stat),2)
igs.mean <- round(sum(igs$Combined.stat * igs$Combined.N) / sum(igs$Combined.N, na.rm = T),2)
```
Data in DAFI show that deliberate anthropogenic fires occur (where present) at a median and mean rate of `r igs.med` and `r igs.mean` km-2 year-1 respectively. 

```{r}
### Fire Return Period - not annualised
FR <- plyr::rbind.fill(lapply(colnames(rep_FUS)[c(27)], 
                               function (x){summarise.behaviour(temp.rep_fire_use=rep_FUS,temp.est_fire_use=est_FUS,
                                                                type = "Fire", behaviour = x, 
                                                                grouping = c('Case Study ID'), escape.rm = T)}))
ri.mean <- round(sum(FR$Combined.stat * FR$Combined.N) / sum(FR$Combined.N, na.rm = T),2)

### median FR - use all values for more accuracy (not possible for igs using aggregation)
all_fire <- rep_FUS
all_fire$`Fire return period (years)` <- ifelse(is.na(all_fire$`Fire return period (years)`), 
                                                unlist(unbin(est_FUS$`Fire return period (years)`)), 
                                                all_fire$`Fire return period (years)`)

all_fire <- all_fire %>% filter(`Fire type` == 'Human, deliberate' & `Presence / Absence` == 'Presence')

ri.med <- round(median(all_fire$`Fire return period (years)`, na.rm = T),2)
```

Similarly, fire return intervals are typically short, with median of `r round(ri.med,1)` and mean of `r round(ri.mean,1)` years.

### Figure 3
```{r warning=F}
fs_mean_kde <- hist.fire.regime(dat.temp=fs.cs.afr,metric = 'fire size mean', bin_width = -1, actual = F, log_scale=T, regime=F, vertical=T, scale_limits = c(0.1, 10000)) 
#ggsave('fs_mean_kde.pdf', plot=fs_mean_kde, height=14, width=9, units='cm')
fs_mean_kde

fs_mean_box <- box.FS.regime(fs.cs.afr, metric = 'mean', boxtype='box', 
                   estimated = T, reported = T, log_scale = T, cropland = T, scale_limits = c(0.1, 10000))
#ggsave('fs_mean_box.pdf', plot=fs_mean_box, height=14, width=16, units='cm')
fs_mean_box 
```


```{r}
ba_ha_box <- box.BA.regime(dat.temp=ba.cs.afr,metric = 'burned area (ha)', boxtype='box', 
              log_scale = T, cropland = T, scale_limits = c(0.01, 1000000000)) 
#ggsave('ba_ha_box.pdf', plot=ba_ha_box, height=14, width=16, units='cm')
ba_ha_box

ba_ha_kde <- hist.fire.regime(dat.temp=ba.cs.afr,metric = 'burned area (ha)', bin_width = -1, log_scale=T, regime=F, vertical=T,  scale_limits = c(0.01, 1000000000)) 
#ggsave('ba_ha_kde.pdf', plot=ba_ha_kde, height=14, width=9, units='cm')
ba_ha_kde
```

## 3.3 Fire Suppression

### Figure 4
```{r warning=F}

#suppression at case study level
#
#only count CaseStudy once for each AFR-Suppression-Effort combination
#don't use pivot_longer as that results in double-counting

cleanEffort <- function(supTbl, supLabel){
  
  rsupTbl <- supTbl %>%
    setNames(c("Case","AFR","Effort")) %>%
    drop_na() %>%
    filter(Effort != 'ND') %>%
    mutate(Effort = recode(Effort, `0`='None',`1`='Limited',`2`='Moderate',`3`='Intensive')) %>%
    group_by_all %>%
    summarise(n= n()) %>%
    dplyr::select(-n) %>%
    mutate(Suppression = supLabel)
  
  return(rsupTbl)
  
}

sup.c <- cleanEffort(dplyr::select(suppression, 1,2,9),'Control')
sup.p <- cleanEffort(dplyr::select(suppression, 1,2,11),'Prevention')
sup.e <- cleanEffort(dplyr::select(suppression, 1,2,13),'Extinction')

c.sup <- rbind(sup.c, sup.p, sup.e)  

c.sup.c <- c.sup %>% 
  ungroup() %>%
  mutate(Effort = factor(Effort, levels = c("None","Limited","Moderate","Intensive"))) %>%
  mutate(AFR = factor(AFR, levels = c("Pre-industrial","Transition","Industrial","Post-industrial"))) %>% 
  mutate(Suppression = factor(Suppression,levels = c("Control","Prevention","Extinction"))) %>%
  count(AFR, Suppression, Effort,.drop=FALSE)

#dodge
fig4 <- c.sup.c %>%
  ggplot(aes(x = AFR, y=n)) +
  geom_bar(aes(fill=Effort),position='dodge',stat="identity") +
  facet_grid(Suppression~.) +
  scale_fill_viridis_d(option = 'inferno') +
  xlab("Anthropogenic Fire Regime") +
  ylab("Count") +
  theme(axis.text.y=element_text(color = "black", size=9),
        axis.text.x=element_text(color = "black", size=9))
#ggsave(fig4, filename='Figure4_600.png', width=14, height=8.5, units='cm', dpi=600)

#fill
fig4_alt <- c.sup.c %>%
  ggplot(aes(x = AFR, y=n)) +
  geom_bar(aes(fill=Effort),position='fill',stat="identity") +
  facet_grid(Suppression~.) +
  scale_fill_viridis_d(option = 'inferno') +
  xlab("Anthropogenic Fire Regime") +
  ylab("Count")

fig4
fig4_alt
```


```{r eval=F, echo=F}
#not used, but retained for reference
c.sup.c %>%
 filter(Effort != "None") %>%
 group_by(AFR, Suppression) %>%
 summarise(sum = sum(n))

c.sup.c %>%
 group_by(AFR, Effort) %>%
 summarise(sum = sum(n))

filter(c.sup, AFR=='Post-industrial', Effort=="Intensive", Suppression=="Extinction")
```

## 3.4 Fire Policy

```{r echo=F}
#figure not included but code retained
pol.sum <- plyr::rbind.fill(lapply(colnames(policy)[c(10,14,19)],
                                     function (x){summarise.behaviour(type = "Policy", behaviour = x,
                                                                      grouping = c('Anthropogenic fire regime','Direct Fire policy ID'),
                                                                      inc.Absence = F, escape.rm = T)}))
pol.sum <-pol.sum[!is.na(pol.sum$`Anthropogenic fire regime`),]
pol.sum <- pol.sum %>% 
    count(`Anthropogenic fire regime`, Behaviour,.drop=FALSE)

pol.sum$`Anthropogenic fire regime` <- factor(pol.sum$`Anthropogenic fire regime`, levels = c("Pre-industrial","Transition","Industrial","Post-industrial"))

p7 <- pol.sum %>%
  ggplot(aes(x = `Anthropogenic fire regime`, y=n)) +
  geom_bar(aes(fill=Behaviour),position='dodge',stat="identity") +
  scale_fill_viridis_d(option = 'inferno') +
  xlab("Anthropogenic Fire Regime") +
  ylab("Count")
```

### Figure 5
```{r}
#policy at case study level 
pol.cs <- policy %>%
  dplyr::select(1,2,4,10:17,19:22) %>%
  rename('Case'=1,'AFR' = 2) %>%
  mutate_at(vars(-1,-2,-3), funs(case_when(grepl("Yes", ., ignore.case=TRUE) ~ "Yes"))) %>%
  dplyr::select(-`Fire type`)

#create 'Other' Rationale for when Type = Yes but individual rationale is not specified
pol.cs <- pol.cs %>% 
  rename('Unknown incentives' = 'Incentives') %>%
  rename('Unknown restriction' = 'Fire restricted') %>%
  rename('Unknown ban' = 'Fire banned') %>%
  mutate(`Unknown incentives` = if_else(`Unknown incentives` == "Yes" & is.na(`Economic incentives`) &
                                          is.na(`Environmental incentives`) & is.na(`Health incentives`), "Yes", NA_character_)) %>%
  mutate(`Unknown restriction` = if_else(`Unknown restriction` == "Yes" & is.na(`Economic restriction`) &
                                          is.na(`Environmental restriction`) & is.na(`Health restriction`), "Yes", NA_character_)) %>%
  mutate(`Unknown ban` = if_else(`Unknown ban` == "Yes" & is.na(`Economic ban`) &
                                          is.na(`Environmental ban`) & is.na(`Health ban`), "Yes", NA_character_))

pol.cs <- pol.cs %>%
  group_by_all %>%
  summarise(n= n()) %>%
  ungroup()

#go long
pol.cs <-pivot_longer(pol.cs, cols=c(`Unknown incentives`:`Health ban`), names_to="Rationale-Type")

#create new columns for Rationale and Type then set factors and levels
pol.cs <- pol.cs %>%
  separate(`Rationale-Type`, c("Rationale", "Type"), sep=" ") %>%
  mutate(Type = recode_factor(Type, incentives='Incentive',restriction='Restriction',ban='Ban')) %>%
  mutate(Rationale = factor(Rationale, levels=c("Environmental","Economic","Health","Unknown"))) %>%
  mutate(AFR = factor(AFR, levels = c("Pre-industrial","Transition","Industrial","Post-industrial")))

#drop NAs and count         
pol.cs <- pol.cs %>% 
  drop_na(value) %>%
  drop_na(AFR) %>%
  count(AFR, Type, Rationale,.drop=FALSE)

#https://stackoverflow.com/a/12104207
rationale_names <- list(
  'Environmental'='Environment',
  'Economic'='Economic',
  'Health'='Health',
  'Unknown'='Unknown'
)

rationale_labeller <- function(variable,value){
  return(rationale_names[value])
}

#dodge
fig5 <- pol.cs %>%
  ggplot(aes(x = AFR, y=n)) +
  geom_bar(aes(fill=Type),position='dodge',stat="identity") +
  facet_grid(Rationale~., labeller=rationale_labeller) +
  scale_fill_viridis_d(option = 'inferno') +
  xlab("Anthropogenic Fire Regime") +
  ylab("Count") +
  theme(axis.text.y=element_text(color = "black", size=9),
        axis.text.x=element_text(color = "black", size=9),
        strip.text = element_text(size = 8))
ggsave(fig5, filename='Figure5_600.png', width=14, height=8.5, units='cm', dpi=600)

#plot stacked
fig5_alt <- pol.cs %>%
  ggplot(aes(x = AFR, y=n)) +
  geom_bar(aes(fill=Type),position='fill',stat="identity") +
  facet_grid(Rationale~.) +
  scale_fill_viridis_d(option = 'inferno') +
  xlab("Anthropogenic Fire Regime") +
  ylab("Count") 

fig5
fig5_alt
```

```{r eval=F, echo=F}
#not used, but retained for reference
pol.cs %>%
  group_by(Type) %>%
  summarise(sum = sum(n))
```
# 4. Discussion

## 4.1 Improving the quality of anthropogenic fire data
```{r}
ba.rep <- ba.cs.afr %>%
      drop_na(`Anthropogenic fire regime`) %>%
      filter(Combined.stat > 0.01) %>%
      filter(grepl('burned area (ha)',Behaviour,fixed=TRUE)) 

fs.rep  <- fs.cs.afr %>%
      drop_na(`Anthropogenic fire regime`) %>%
      filter(Combined.stat > 0.01) %>%
      filter(grepl('mean',Behaviour,fixed=TRUE))

ba.rep <- round(100 * length(ba.rep[,1]) / n_distinct(record_info$`Case.Study.ID`),1)
fs.rep <- round(100 * length(fs.rep[,1]) / n_distinct(record_info$`Case.Study.ID`),1)
```

we find that in case studies, burned area metrics are reported less frequently than fire size metrics (`r ba.rep`% of case studies vs `r fs.rep`%). 

##  4.2 Modelling and observing anthropogenic fire regimes 

### Figure 6
```{r warning=F}
#https://datascience.blog.wzb.eu/2019/04/30/zooming-in-on-maps-with-sf-and-ggplot2/
##https://ggplot2.tidyverse.org/reference/ggsf.html
require(sf)
require(rnaturalearth)
require(rnaturalearthdata)
worldmap <- ne_countries(scale = 'medium', type = 'map_units',returnclass = 'sf')

dat <- right_join(record_info, est_fire_use, by=c("Case.Study.ID"='Case Study ID'))

dat$Latitude <- as.numeric(dat$Latitude)
dat$Longitude<- as.numeric(dat$Longitude) 

fsdat <- dat %>% 
  filter(grepl('Crop',`Fire purpose`)) %>%
  drop_na(`Fire season start`) %>%
  mutate(`Fire season start` = factor(`Fire season start`, levels = month.name)) %>%
  mutate(`Fire quarter` = factor(case_when(
    `Fire season start` == 'December' | `Fire season start` == 'January' | `Fire season start` == 'February' ~ 'DJF',
    `Fire season start` == 'March' | `Fire season start` == 'April' | `Fire season start` == 'May' ~ 'MAM',
    `Fire season start` == 'June' | `Fire season start` == 'July' | `Fire season start` == 'August' ~ 'JJA',
    `Fire season start` == 'September' | `Fire season start` == 'October' | `Fire season start` == 'November' ~ 'SON'), 
    levels= c('DJF','MAM','JJA','SON')))

shapes = c(21,23) 
inset_x <- c(70, 95)
inset_y <- c(20, 33)

#hack described here: https://github.com/tidyverse/ggplot2/issues/3909#issuecomment-602253881
myrect <- 
  list(
    cbind(
      c(inset_x[1], inset_x[2], inset_x[2], inset_x[1], inset_x[1]), 
      c(inset_y[1], inset_y[1], inset_y[2], inset_y[2], inset_y[1])
    )
  ) %>%
  st_polygon() %>%
  st_sfc(crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

#world plot
w <- ggplot() +
  geom_sf(data = worldmap, fill="beige", colour='grey',size=0.1) +
  geom_sf(data = myrect,
  #geom_rect(xmin = 50, xmax = 70, ymin = 50, ymax = 70, 
     fill = NA, colour = "red", size = 0.5) +
  coord_sf(xlim = c(-170, 180), ylim = c(-60, 85), expand = FALSE) +
  geom_point(data=fsdat,aes(x = Longitude, y = Latitude, shape = `Fire purpose`, fill= `Fire quarter`),
             colour='black', size = 0.75, stroke=0.1, alpha=0.9) + 
  theme(
    panel.background = element_rect(fill = "azure",colour = "black",size = 0.5, linetype = "solid"),
    panel.grid.major = element_line(linetype='blank'),
    panel.grid.minor = element_line(linetype='blank'),
    legend.key = element_rect(fill = "white", colour = "white")
  ) +
  scale_fill_viridis_d(option='inferno') + 
  guides(fill=guide_legend(override.aes=list(shape=22))) +
  scale_shape_manual(values=shapes) 

#north India inset
i <- ggplot() + 
  geom_sf(data = worldmap, fill='beige', colour='grey') + 
  coord_sf(xlim = inset_x, ylim = inset_y, expand = FALSE) +
  geom_point(data=fsdat, aes(x = Longitude, y = Latitude, shape = `Fire purpose`, fill= `Fire quarter`),
             colour='black', size = 2, stroke=0.5, alpha=0.9) +
    theme(
    panel.background = element_rect(fill = "azure",colour = "red",size = 0.5, linetype = "solid"),
    panel.grid.major = element_line(linetype='blank'),
    panel.grid.minor = element_line(linetype='blank'),
    legend.key = element_rect(fill = "white", colour = "white")
  ) +
  scale_fill_viridis_d(option='inferno') + 
  guides(fill=guide_legend(override.aes=list(shape=22))) +
  scale_shape_manual(values=shapes) 
```

```{r warning=F}
w
i
```

```{r}
count21ha <- rep_FUS %>%
  filter(`Actual fire size mean (ha)` > 21) %>%
  filter(`Presence / Absence` == 'Presence') %>%
  filter(`Fire type` == 'Human, deliberate') %>%
  nrow(.)
  
count0ha <- rep_FUS %>%
  filter(`Actual fire size mean (ha)` > 0) %>%
  filter(`Presence / Absence` == 'Presence') %>%
  filter(`Fire type` == 'Human, deliberate') %>%
  nrow(.)  

perc21 <- 100 - (100 * as.numeric(count21ha) / as.numeric(count0ha))

count100ha <- rep_FUS %>%
  filter(`Actual fire size mean (ha)` > 100) %>%
  filter(`Presence / Absence` == 'Presence') %>%
  filter(`Fire type` == 'Human, deliberate') %>%
  nrow(.)
  
perc100 <- 100 - (100 * as.numeric(count100ha) / as.numeric(count0ha))
```

We find that `r round(perc21,0)`% of records in DAFI for mean size of deliberately started fires are <21 ha, suggesting many anthropogenic fires will not be detected. 

Use of Sentinel‐2 derived fire data (at 0.04 ha spatial resolution) has shown how MODIS‐derived products may underestimate burned area by up to 80% across Africa and that fires <100 ha are critically important for characterizing landscape fire on a global scale [102]. Our results from DAFI support this; `r round(perc100,0)`% of mean fire size records for deliberately started fires are <100ha, and cropland fires in pre-industrial and transition regimes are generally smaller compared to fires ignited to burn across landscapes more broadly (Figure 3a).

We also find a discrepancy between the median density for deliberate anthropogenic fires in DAFI of `r igs.med` (Section 3.2.2) and the median value suggested by the MODIS‐derived global fire atlas of less than 0.01 fires km−2 year−1 [44]. The mean density in DAFI of `r igs.mean` km−2 year−1 indicates a skewed distribution and the DAFI‐MODIS discrepancy is most acute in regions of intensive crop residue burning in tightly packed fields.

## 4.3. Categorising Anthropogenic Fire Uses and Regimes

```{r}
fs.fp <- plyr::rbind.fill(lapply(colnames(rep_fire_use)[13:20],
                                     function (x){summarise.behaviour(temp.rep_fire_use=rep_FUS,temp.est_fire_use=est_FUS,
                                                                      type = "Fire", behaviour = x,
                                                                      grouping = c('Fire purpose'),
                                                                      inc.Absence = F, escape.rm = T)})) %>%
  filter(`Fire purpose` %in% c('Hunter gatherer', 'Pasture management')) %>%
  filter((grepl('max', Behaviour) | grepl('min', Behaviour)) & grepl('Actual', Behaviour))

hg.min <- filter(fs.fp, `Fire purpose` == 'Hunter gatherer' & grepl('min', Behaviour))$Combined.stat
hg.max <- filter(fs.fp, `Fire purpose` == 'Hunter gatherer' & grepl('max', Behaviour))$Combined.stat

pas.min <- filter(fs.fp, `Fire purpose` == 'Pasture management' & grepl('min', Behaviour))$Combined.stat
pas.max <- filter(fs.fp, `Fire purpose` == 'Pasture management' & grepl('max', Behaviour))$Combined.stat

```
In particular, hunting-gathering fire sizes vary more than pasture managament fires, with sizes ranging from `r round(hg.min,1)` ha to `r round(hg.max,1)` ha compared to `r round(pas.min,1)` ha to `r round(pas.max,1)` ha.
